---
title: "Integration_human_otherMethods"
description: |
  Overview of data analysis.
author:
  - name: Manuela Benary / Mohammad El-Ismail
    affiliation: CUBI / CCCC
    affiliation_url: 
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

```{r load-libraries}
library(Seurat)
library(tidyverse)
library(RCurl)
library(dplyr)
library(cowplot)
library(ggsci)
```



## Fast Integration using RPCA
#split_filtered cond seurat without regression

```{r}
split_filtered_cond_seurat <- readRDS("E://CCCC/PI3K/data/SeuratObjects/splitted/Human/split_filtered_cond_seurat.rds")
```

```{r}
split_filtered_cond_seurat <- lapply(X = split_filtered_cond_seurat, FUN = function(x) {
     x <- NormalizeData(x)
     x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
 })
```

```{r}
features <- SelectIntegrationFeatures(object.list = split_filtered_cond_seurat)
split_filtered_cond_seurat <- lapply(X = split_filtered_cond_seurat, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
   x <- RunPCA(x, features = features, verbose = FALSE)
 })
```

```{r}
integrated_filtered_rpca_cond_seurat <- FindIntegrationAnchors(object.list = split_filtered_cond_seurat, anchor.features = features, reduction = "rpca")
```

```{r}
integrated_filtered_rpca_cond_seurat <- IntegrateData(anchorset = integrated_filtered_rpca_cond_seurat)
```

```{r}
# Run the standard workflow for visualization and clustering
integrated_filtered_rpca_cond_seurat <- ScaleData(integrated_filtered_rpca_cond_seurat, verbose = FALSE)
integrated_filtered_rpca_cond_seurat <- RunPCA(integrated_filtered_rpca_cond_seurat, npcs = 30, verbose = FALSE)
integrated_filtered_rpca_cond_seurat <- RunUMAP(integrated_filtered_rpca_cond_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_cond_seurat <- FindNeighbors(integrated_filtered_rpca_cond_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_cond_seurat <- FindClusters(integrated_filtered_rpca_cond_seurat, resolution = 0.5)
```   

```{r}
p1_rpca <- DimPlot(integrated_filtered_rpca_cond_seurat, reduction = "umap", group.by = "Sample_Tag") +
      labs(title = "UMAP visualisation by conditions without regression") +
      scale_fill_lancet()
p2_rpca <- DimPlot(integrated_filtered_rpca_cond_seurat, reduction = "umap", label = TRUE, repel = TRUE)
p3_rpca <- DimPlot(integrated_filtered_rpca_cond_seurat, reduction = "umap", split.by = "Sample_Tag")
p1_rpca
p2_rpca + scale_fill_lancet()
p3_rpca + scale_fill_lancet()
```

```{r}
saveRDS(integrated_filtered_rpca_cond_seurat, "add path were you want to save the seuratobject")
```

## Performing integration on datasets normalized with SCTransform
# filtered_sct_mt_cond_seurat with mitochondrial gene expression regression

```{r load-data}
split_filtered_sct_mt_cond_seurat <- readRDS("add path where you saved the seuratobject (filtered_sct_mt_cond_seurat) in file /PI3K/Normalization/otherMethods/seurat_norm_cond_otherMethods_human.Rmd")
```

```{r }
integ_features <- SelectIntegrationFeatures(object.list = split_filtered_sct_mt_cond_seurat,
                                          nfeatures = 3000)
```

```{r select features and run PCA}
split_filtered_sct_mt_cond_seurat <- PrepSCTIntegration(object.list = split_filtered_sct_mt_cond_seurat,
                                 anchor.features = integ_features)
```

```{r}
integ_anchors <- FindIntegrationAnchors(object.list = split_filtered_sct_mt_cond_seurat,
                                       normalization.method = "SCT",
                                       anchor.features = integ_features)
```

```{r Integrate across conditions}
integrated_sct_mt_cond_seurat <- IntegrateData(anchorset = integ_anchors,
                                         normalization.method = "SCT")
```

```{r downstream analysis on the corrected data}
DefaultAssay(integrated_sct_mt_cond_seurat) <- "integrated"
```

```{r workflow for visualization and clustering}
integrated_sct_mt_cond_seurat <- ScaleData(integrated_sct_mt_cond_seurat, verbose = FALSE)
integrated_sct_mt_cond_seurat <- RunPCA(integrated_sct_mt_cond_seurat, npcs = 30, verbose = FALSE)
integrated_sct_mt_cond_seurat <- RunUMAP(integrated_sct_mt_cond_seurat, reduction = "pca", dims = 1:30)
integrated_sct_mt_cond_seurat <- FindNeighbors(integrated_sct_mt_cond_seurat, reduction = "pca", dims = 1:30)
integrated_sct_mt_cond_seurat <- FindClusters(integrated_sct_mt_cond_seurat, resolution = 0.5)
```

```{r Visualization}
p4 <- DimPlot(integrated_sct_mt_cond_seurat, reduction = "umap", group.by = "Sample_Tag")
p5 <- DimPlot(integrated_sct_mt_cond_seurat, reduction = "umap", label = TRUE, repel = TRUE)
p6 <- DimPlot(integrated_sct_mt_cond_seurat, reduction = "umap", split.by = "Sample_Tag")
p4 + scale_fill_lancet()
p5 + scale_fill_lancet()
p6 + scale_fill_lancet()
```

```{r}
saveRDS(integrated_sct_mt_cond_seurat, "add path were you want to save the seuratobject")
```


## Fast Integration using RPCA
#split_filtered_cond seurat with mitochondrial gene expression regression

```{r}
split_filtered_cond_seurat <- readRDS("E://CCCC/PI3K/data/SeuratObjects/splitted/Human/split_filtered_cond_seurat.rds")
```

```{r}
split_filtered_cond_seurat <- lapply(X = split_filtered_cond_seurat, FUN = function(x) {
   x <- NormalizeData(x)
   x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
```

```{r}
features <- SelectIntegrationFeatures(object.list = split_filtered_cond_seurat)
split_filtered_seurat <- lapply(X = split_filtered_cond_seurat, FUN = function(x) {
   x <- ScaleData(x, vars.to.regress = c("percent.mt"), features = features, verbose = FALSE)
   x <- RunPCA(x, features = features, verbose = FALSE)
})
```

```{r}
integrated_filtered_rpca_mt_cond_seurat <- FindIntegrationAnchors(object.list = split_filtered_cond_seurat, anchor.features = features, reduction = "rpca")
```

```{r}
integrated_filtered_rpca_mt_cond_seurat <- IntegrateData(anchorset = integrated_filtered_rpca_mt_cond_seurat)
```

```{r}
#Run the standard workflow for visualization and clustering
integrated_filtered_rpca_mt_cond_seurat <- ScaleData(integrated_filtered_rpca_mt_cond_seurat, verbose = FALSE)
integrated_filtered_rpca_mt_cond_seurat <- RunPCA(integrated_filtered_rpca_mt_cond_seurat, npcs = 30, verbose = FALSE)
integrated_filtered_rpca_mt_cond_seurat <- RunUMAP(integrated_filtered_rpca_mt_cond_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_mt_cond_seurat <- FindNeighbors(integrated_filtered_rpca_mt_cond_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_mt_cond_seurat <- FindClusters(integrated_filtered_rpca_mt_cond_seurat, resolution = 0.5)
```   

```{r}
p4_rpca <- DimPlot(integrated_filtered_rpca_mt_cond_seurat, reduction = "umap", group.by = "Xenograft")
p5_rpca <- DimPlot(integrated_filtered_rpca_mt_cond_seurat, reduction = "umap", label = TRUE, repel = TRUE)
p6_rpca <- DimPlot(integrated_filtered_rpca_mt_cond_seurat, reduction = "umap", split.by = "Xenograft")
p4_rpca + scale_fill_lancet()
p5_rpca + scale_fill_lancet()
p6_rpca + scale_fill_lancet()
```

```{r}
saveRDS(integrated_filtered_rpca_mt_cond_seurat, "add path were you want to save the seuratobject")
```



## Performing integration on datasets normalized with SCTransform
# filtered_sct_phase_cond_seurat with cell cycle phase regression

```{r}
split_filtered_sct_phase_cond_seurat <- readRDS("add path where you saved the seuratobject (filtered_sct_phase_cond_seurat) in file /PI3K/Normalization/otherMethods/seurat_norm_cond_otherMethods_human.Rmd")
```

```{r }
integ_features <- SelectIntegrationFeatures(object.list = split_filtered_sct_phase_cond_seurat,
                                            nfeatures = 3000)
```

```{r}
split_filtered_sct_phase_cond_seurat <- PrepSCTIntegration(object.list = split_filtered_sct_phase_cond_seurat,
                                   anchor.features = integ_features)
```

```{r}
integ_anchors <- FindIntegrationAnchors(object.list = split_filtered_sct_phase_cond_seurat,
                                        normalization.method = "SCT",
                                        anchor.features = integ_features)
```

```{r}
integrated_sct_phase_cond_seurat <- IntegrateData(anchorset = integ_anchors,
                                          normalization.method = "SCT")
```

```{r}
DefaultAssay(integrated_sct_phase_seurat) <- "integrated"
```

```{r}
integrated_sct_phase_cond_seurat <- ScaleData(integrated_sct_phase_cond_seurat, verbose = FALSE)
integrated_sct_phase_cond_seurat <- RunPCA(integrated_sct_phase_cond_seurat, npcs = 30, verbose = FALSE)
integrated_sct_phase_cond_seurat <- RunUMAP(integrated_sct_phase_cond_seurat, reduction = "pca", dims = 1:30)
integrated_sct_phase_cond_seurat <- FindNeighbors(integrated_sct_phase_cond_seurat, reduction = "pca", dims = 1:30)
integrated_sct_phase_cond_seurat <- FindClusters(integrated_sct_phase_cond_seurat, resolution = 0.5)
```


```{r}
p7 <- DimPlot(integrated_sct_phase_cond_seurat, reduction = "umap", group.by = "Xenograft")
p8 <- DimPlot(integrated_sct_phase_cond_seurat, reduction = "umap", label = TRUE, repel = TRUE)
p9 <- DimPlot(integrated_sct_phase_cond_seurat, reduction = "umap", split.by = "Sample_Tag")
p7 + scale_fill_lancet()
p8 + scale_fill_lancet()
p9 + scale_fill_lancet()
```

```{r}
saveRDS(integrated_sct_phase_cond_seurat, "add path were you want to save the seuratobject")
```


## Fast Integration using RPCA
#split_filtered_seurat with cell cycle phase regression

```{r}
split_filtered_seurat <- readRDS("E://CCCC/PI3K/data/SeuratObjects/splitted/Human/split_filtered_cond_seurat.rds")
```

```{r}
split_filtered_seurat <- lapply(X = split_filtered_seurat, FUN = function(x) {
   x <- NormalizeData(x)
   x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
```

```{r}
features <- SelectIntegrationFeatures(object.list = split_filtered_seurat)
split_filtered_seurat <- lapply(X = split_filtered_seurat, FUN = function(x) {
   x <- ScaleData(x, vars.to.regress = c("S.Score", "G2M.Score"), features = features, verbose = FALSE)
   x <- RunPCA(x, features = features, verbose = FALSE)
})
```

```{r}
integrated_filtered_rpca_phase_seurat <- FindIntegrationAnchors(object.list = split_filtered_seurat, anchor.features = features, reduction = "rpca")
```

```{r}
integrated_filtered_rpca_phase_seurat <- IntegrateData(anchorset = integrated_filtered_rpca_phase_seurat)
```

```{r}
# Run the standard workflow for visualization and clustering
integrated_filtered_rpca_phase_seurat <- ScaleData(integrated_filtered_rpca_phase_seurat, verbose = FALSE)
integrated_filtered_rpca_phase_seurat <- RunPCA(integrated_filtered_rpca_phase_seurat, npcs = 30, verbose = FALSE)
integrated_filtered_rpca_phase_seurat <- RunUMAP(integrated_filtered_rpca_phase_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_phase_seurat <- FindNeighbors(integrated_filtered_rpca_phase_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_phase_seurat <- FindClusters(integrated_filtered_rpca_phase_seurat, resolution = 0.5)
```   

```{r}
p7_rpca <- DimPlot(integrated_filtered_rpca_phase_seurat, reduction = "umap", group.by = "Xenograft")
p8_rpca <- DimPlot(integrated_filtered_rpca_phase_seurat, reduction = "umap", label = TRUE, repel = TRUE)
p9_rpca <- DimPlot(integrated_filtered_rpca_phase_seurat, reduction = "umap", split.by = "Sample_Tag")
p7_rpca
p8_rpca
p9_rpca
```

```{r}
saveRDS(integrated_filtered_rpca_phase_seurat, "add path were you want to save the seuratobject")
```



## Performing integration on datasets normalized with SCTransform
# filtered_sct_mt_phase_seurat with mitochondrial gene expression and cell cycle phase regression

```{r}
split_filtered_sct_mt_phase_seurat <- readRDS("add path where you saved the seuratobject (filtered_sct_mt_phase_seurat) in file /PI3K/Normalization/otherMethods/seurat_norm_cond_otherMethods_human.Rmd")
```

```{r }
integ_features <- SelectIntegrationFeatures(object.list = split_filtered_sct_mt_phase_seurat,
                                            nfeatures = 3000)
```

```{r}
split_filtered_sct_mt_phase_seurat <- PrepSCTIntegration(object.list = split_filtered_sct_mt_phase_seurat, anchor.features = integ_features)
```

```{r}
integ_anchors <- FindIntegrationAnchors(object.list = split_filtered_sct_mt_phase_seurat, normalization.method = "SCT", anchor.features = integ_features)
```

```{r}
integrated_sct_mt_phase_seurat <- IntegrateData(anchorset = integ_anchors,
                                         normalization.method = "SCT")
```

```{r}
DefaultAssay(integrated_sct_mt_phase_seurat) <- "integrated"
```

```{r}
integrated_sct_mt_phase_seurat <- ScaleData(integrated_sct_mt_phase_seurat, verbose = FALSE)
integrated_sct_mt_phase_seurat <- RunPCA(integrated_sct_mt_phase_seurat, npcs = 30, verbose = FALSE)
integrated_sct_mt_phase_seurat <- RunUMAP(integrated_sct_mt_phase_seurat, reduction = "pca", dims = 1:30)
integrated_sct_mt_phase_seurat <- FindNeighbors(integrated_sct_mt_phase_seurat, reduction = "pca", dims = 1:30)
integrated_sct_mt_phase_seurat <- FindClusters(integrated_sct_mt_phase_seurat, resolution = 0.5)
```

```{r}
p10 <- DimPlot(integrated_sct_mt_phase_seurat, reduction = "umap", group.by = "Sample_Tag")
p11 <- DimPlot(integrated_sct_mt_phase_seurat, reduction = "umap", label = TRUE, repel = TRUE)
p12 <- DimPlot(integrated_sct_mt_phase_seurat, reduction = "umap", split.by = "Sample_Tag")
p10 +  scale_fill_lancet()
p11
p12
```

```{r}
saveRDS(integrated_sct_mt_phase_seurat, "add path were you want to save the seuratobject")
```


## Fast Integration using RPCA
#split_filtered_seurat with mt and phase regression

```{r}
split_filtered_seurat <- readRDS("E://CCCC/PI3K/data/SeuratObjects/splitted/Human/split_filtered_cond_seurat.rds")
```

```{r}
split_filtered_seurat <- lapply(X = split_filtered_seurat, FUN = function(x) {
   x <- NormalizeData(x)
   x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
```

```{r}
features <- SelectIntegrationFeatures(object.list = split_filtered_seurat)
split_filtered_seurat <- lapply(X = split_filtered_seurat, FUN = function(x) {
   x <- ScaleData(x, vars.to.regress = c("S.Score", "G2M.Score", "percent.mt"), features = features, verbose = FALSE)
   x <- RunPCA(x, features = features, verbose = FALSE)
})
```

```{r}
integrated_filtered_rpca_mt_phase_seurat <- FindIntegrationAnchors(object.list = split_filtered_seurat, anchor.features = features, reduction = "rpca")
```

```{r}
integrated_filtered_rpca_mt_phase_seurat <- IntegrateData(anchorset = integrated_filtered_rpca_mt_phase_seurat)
```

```{r}
# Run the standard workflow for visualization and clustering
integrated_filtered_rpca_mt_phase_seurat <- ScaleData(integrated_filtered_rpca_mt_phase_seurat, verbose = FALSE)
integrated_filtered_rpca_mt_phase_seurat <- RunPCA(integrated_filtered_rpca_mt_phase_seurat, npcs = 30, verbose = FALSE)
integrated_filtered_rpca_mt_phase_seurat <- RunUMAP(integrated_filtered_rpca_mt_phase_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_mt_phase_seurat <- FindNeighbors(integrated_filtered_rpca_mt_phase_seurat, reduction = "pca", dims = 1:30)
integrated_filtered_rpca_mt_phase_seurat <- FindClusters(integrated_filtered_rpca_mt_phase_seurat, resolution = 0.5)
```   

```{r}
p10_rpca <- DimPlot(integrated_filtered_rpca_mt_phase_seurat, reduction = "umap", group.by = "Sample_Tag")
p11_rpca <- DimPlot(integrated_filtered_rpca_mt_phase_seurat, reduction = "umap", label = TRUE, repel = TRUE)
p12_rpca <- DimPlot(integrated_filtered_rpca_mt_phase_seurat, reduction = "umap", split.by = "Sample_Tag")
p10_rpca + scale_fill_lancet()
p11_rpca
p12_rpca
```

```{r}
saveRDS(integrated_filtered_rpca_mt_phase_seurat, "add path were you want to save the seuratobject")
```